---
title: "Transportation & Emissions"
author: "Ali Sanford"
date: "May 1, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(DataComputing)
library(dplyr)
library(geosphere) #contains the distance function that we used
library(ggplot2)
library(tidyr)
library(statisticalModeling)
library(RColorBrewer)
library(ggmap)
knitr::opts_chunk$set(echo = TRUE)
```

## The Relationship Between Public Transportation and Emissions in Two Major CA Cities
<div class="blue3">Is a neighborhood's proximity to public transit stops related to its carbon emissions?</div>

<div class="green3">-We hypothesized that zip codes with more public transit stops would have less carbon emissions overall</div>
<div class="green3">-We thought it would be interesting to see if this relationship was affected by a zip code's average household income</div>
<div class="green3">-We also wondered if the relationship changes when carbon emissions are adjusted for the population of the zip code</div>


## Data Sets and Methods {.smaller}
<div class="blue3">Emissions Data</div>
<div class="green3">-The Cool Climate Network provided us with a dataset including carbon emissions for every zip code in the US, broken down by emission type</div>
<div class="green3">-It also included other information about each zip code</div>
<div class="green3">-We filtered the data for zip codes within Los Angeles and San Francisco, and selected only the relevant variables, including Longitude and Latitude, Transport Emissions and Average Income Per Household</div>
<div class="blue3">Transit Stop Data</div>
<div class="green3">-Public transit systems make their system data available in GTFS format, which includes .txt files with each stop's name, longitude and latitude</div>
<div class="green3">-We joined together the data for all transit systems in LA and in SF so that we had the coordinates for each stop in each city</div>
<div class="blue3">Methods</div>
<div class="green3">-We cleaned the data using functions such as `mutate()`, `filter()`, `select()`, `join()`, and others that we learned in class</div>
<div class="green3">-We calculated distances between zip codes and transit stops and determined the number of stops nearby each zip code using nested for loops with if/else statements inside them</div>
<div class="green3">-We also brought in things from outside of class - the geo.sphere package, arcGIS mapping, and ggmapping</div>



## Visualizing the Relationship
```{r include = FALSE}
LAnarrowslides <- read.csv("/Users/alisanford/Desktop/statsfinalproject/statsquad/LAnarrow.csv")
SFnarrowslides <- read.csv("/Users/alisanford/Desktop/statsfinalproject/statsquad/SFnarrow.csv")
emissionsdataLA1slides <- read.csv("/Users/alisanford/Desktop/statsfinalproject/statsquad/emissionsdataLA1.csv")
emissionsdataSF1slides <- read.csv("/Users/alisanford/Desktop/statsfinalproject/statsquad/emissionsdataSF1.csv")

```


```{r}
laplotfaceted <- LAnarrowslides %>%
  ggplot(aes(x = number_of_stops, y = Transport.tCO2e.yr., col = distance)) + 
  geom_point(alpha = .4) + 
  stat_smooth(method = "loess") + facet_grid(. ~ distance, switch = "both") +
  labs(title="Carbon Emissions vs Public Transit Stops - Los Angeles", x="Number of stops", y="CO2 emissions due to Transportation (tons/year)")
laplotfaceted
```




```{r}
sfplotfaceted <- SFnarrowslides %>%
  ggplot(aes(x = number_of_stops, y = Transport.tCO2e.yr., col = distance)) + geom_point(alpha = .4) + stat_smooth(method = "loess") + facet_grid(. ~ distance, switch = "both") +
  labs(title="Carbon Emissions vs Public Transit Stops - San Francisco", x="Number of stops", y="CO2 emissions due to Transportation (tons/year)")
sfplotfaceted
```



```{r include = FALSE}
emissionsdataLA1slides$stoplevels <- cut(emissionsdataLA1slides$within1km, 
                       breaks = c(-Inf, 25, 50, 75, 100, 125, 150, 175, 200, Inf), 
                       labels = c("0-25", "26-50", "51-75", "76-100", "101-125", "126-150", "151-175", "176-200", "201+"), 
                       right = FALSE)

emissionsdataSF1slides$stoplevels <- cut(emissionsdataSF1slides$within1km, 
                       breaks = c(-Inf, 25, 50, 75, 100, 125, 150, 175, 200, Inf), 
                       labels = c("0-25", "26-50", "51-75", "76-100", "101-125", "126-150", "151-175", "176-200", "201+"), 
                       right = FALSE)

emissionsdataLA1slides$emissionlevels <- cut(emissionsdataLA1slides$`Transport.tCO2e.yr.`, 
                       breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 35, 40, Inf), 
                       labels = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40", "41+"), 
                       right = FALSE)

emissionsdataSF1slides$emissionlevels <- cut(emissionsdataSF1slides$`Transport.tCO2e.yr.`, 
                       breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 35, 40, Inf), 
                       labels = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40", "41+"), 
                       right = FALSE)
```

```{r}
ggplot(emissionsdataLA1slides, aes(x = stoplevels, fill = emissionlevels)) +
  geom_bar(position = "fill", alpha = 0.6) +
  scale_fill_brewer(palette = "Spectral") +
  labs(title="Number of Stops vs Emission Levels - LA", x="Number of stops within 1km", legend = "Carbon Emissions from Transportation (tons/yr)")
```


```{r}
ggplot(emissionsdataSF1slides, aes(x = stoplevels, fill = emissionlevels)) +
  geom_bar(position = "fill", alpha = 0.6) +
  scale_fill_brewer(palette = "Spectral") +
  labs(title="Number of Stops vs Emission Levels - SF", x="Number of stops within 1km", legend = "Carbon Emissions from Transportation (tons/yr)")
```




